---
title: "Uncertainty In Cognitive Science"
author: Jesper Fischer Ehmsen,
output:
  word_document:
    reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
  html_document: default
bibliography: [Knitting files/Master_thesis_refs.bib, Knitting files/packages.bib, Knitting files/rpackage.bib]
link-citations: yes
linkcolor: blue
csl: Knitting files/apa.csl
always_allow_html: yes
---


```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")

packages_used = c("tidyverse", "here", "knitr","yaml", "cowplot","flextable", "correlation","cmdstanr","stats")

knit = knitr::opts_knit$get("rmarkdown.pandoc.to")

knitr::write_bib(x = packages_used, file = here::here("manuscript","Knitting files","packages.bib"))

#pacman::p_load(c("tidyverse", "here", "knitr","yaml", "cowplot","flextabe", "correlation"))

lapply(packages_used, library, character.only = TRUE)

```

\newpage

# Summary




\newpage

# Acknowledgement

part of the computation done for this project was performed on the UCloud interactive HPC system, which is managed by the eScience Center at the University of Southern Denmark. 

All scripts, code, models and simulated data can be found on the following [Github page](https://github.com/JesperFischer/Master-thesis). Here all analyses as well as the entire manuscript is documented. 

\newpage

```{r Introduction, child="1_Introduction.Rmd"}

```

\newpage

```{r Measurement uncertainty_v2, child="2_Measurement uncertainty_v2.Rmd"}

```

\newpage

```{r Modeling definitions, child="3_Modeling definitions.Rmd"}

```

\newpage

```{r uncertainty minimazation, child="4_uncertainty minimazation.Rmd"}

```

\newpage

```{r Real data, child="5_Real data.Rmd"}

```

\newpage

```{r Power analysis, child="6_Power analysis.Rmd"}

```

\newpage

```{r Discussion, child="7_Discussion.Rmd"}

```

\newpage

# References


::: {#refs}
:::

\newpage
```{r supplementary, child="Supplementary material/supplementary material.Rmd"}

```
