---
title: "Uncertainty In Cognitive Science"
author: Jesper Fischer Ehmsen,
output:
  word_document:
    reference_docx: Knitting files/docx_template.docx
  pdf_document:
    latex_engine: xelatex
  html_document: default
bibliography: [Knitting files/Master_thesis_refs.bib, Knitting files/packages.bib]
link-citations: yes
linkcolor: blue
csl: Knitting files/apa.csl
always_allow_html: yes
---


```{r load packages, warning = F, message = F, echo = F, include=FALSE}
knitr::opts_chunk$set(dpi=300)

# seed
set.seed(123)

# packages
if (!require("pacman")) install.packages("pacman")

packages_used = c("tidyverse", "here", "knitr","yaml", "cowplot","flextable", "correlation","cmdstanr")

knit = knitr::opts_knit$get("rmarkdown.pandoc.to")

knitr::write_bib(x = packages_used, file = here::here("manuscript","Knitting files","packages.bib"))

#pacman::p_load(c("tidyverse", "here", "knitr","yaml", "cowplot","flextabe", "correlation"))

lapply(packages_used, library, character.only = TRUE)

```

\newpage

# Abstract

Trusting papers introducing models without formally testing them is like trusting a mathematician to measure a 1m stick because of a phd in mathematics or physics instead of having him measure the stick and compare it to the 1m stick. 


\newpage

# Acknowledgement

part of the computation done for this project was performed on the UCloud interactive HPC system, which is managed by the eScience Center at the University of Southern Denmark.

\newpage

```{r Introduction, child="Introduction.Rmd"}

```


```{r Measurement uncertainty, child="Measurement uncertainty.Rmd"}

```


```{r Modeling definitions, child="Modeling definitions.Rmd"}

```


```{r uncertainty minimazation, child="uncertainty minimazation.Rmd"}

```

```{r Real data, child="Real data.Rmd"}

```


```{r Power analysis, child="Power analysis.Rmd"}

```


```{r Discussion, child="Discussion.Rmd"}

```



# References


::: {#refs}
:::

\newpage
