---
title: "ICC psychoRT"
output: html_document
date: "2024-03-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,RWiener, tidybayes, posterior, furrr,gganimate, cmdstanr,patchwork, gamlss,truncnorm,extraDistr,flextable,lmerTest,lme4,Matrix,pracma)
```

## R Markdown

```{r}
source(here::here("realshit","run_ICC_psychoRT.R"))

parameters = expand.grid(subs = c(30),
                         trials = seq(20,100,20),
                         replicate = 1:2) %>% 
  mutate(id = rev(1:nrow(.)))


parameters = parameters %>% mutate(id = 1:nrow(.))

data_list <- split(parameters, parameters$id)



data_list <- split(parameters, parameters$id)

cores = 5
  
plan(multisession, workers = cores)
  
possfit_model = possibly(.f = together, otherwise = "Error")

#test = together(data_list[[1]])

results <- future_map(data_list, ~possfit_model(.x), .progress = TRUE, .options = furrr_options(seed = TRUE,scheduling = 2L))
save.image(here::here("workspace psychometric","ICC_psycho_lowtrials_finished_beta=3 comparison RT_2reps.RData"))


rstudioapi::restartSession()
```



```{r}
source(here::here("realshit","run_ICC_psychoRT.R"))

parameters = expand.grid(subs = c(30),
                         trials = seq(20,100,20),
                         replicate = 1:5) %>% 
  mutate(id = rev(1:nrow(.)))

parameters <- parameters[sample(nrow(parameters)), ]

parameters = parameters %>% mutate(id = 1:nrow(.))

data_list <- split(parameters, parameters$id)

cores = 5
  
plan(multisession, workers = cores)
  
possfit_model = possibly(.f = fit_without_data_gen, otherwise = "Error")

test = together(data_list[[25]])

results <- future_map(data_list, ~possfit_model(.x), .progress = TRUE, .options = furrr_options(seed = TRUE,scheduling = 2L))
save.image(here::here("ICC_psycho_lowtrials_finished_beta=3 comparison RT_real.RData"))


rstudioapi::restartSession()
```



```{r}
load(here::here("workspace psychometric","ICC_psycho_lowtrials_finished_beta=3 rts.RData"))

results <- unlist(results, recursive = FALSE)

results2 = results[-34]


# Step 2: Use map_dfr to row-bind all dataframes
params_icc <- map_dfr(results2, 1)


params_icc %>% ggplot()+
  geom_point(aes(x = trials, y = mean, ymax = q95, ymin = q5, col = interaction(residual_variance, subs)), position = position_jitter(width = 5, height = 0))+
  facet_wrap(~variable)

params_icc %>% filter(grepl("ICC",variable) & residual_variance == T) %>% mutate(subs = as.factor(subs)) %>% 
  ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_wrap(~variable, nrow = 2, scales = "free")+
  geom_point()


params_icc %>% filter(grepl("ICC",variable) & residual_variance == T) %>% group_by(trials,variable) %>% 
  summarize(mean = mean(mean)) %>% 
    ggplot(aes(y = mean, x = trials))+facet_grid(~variable)+
  geom_point()


params_icc %>% filter(grepl("ICC",variable) & residual_variance == T) %>% group_by(subs,trials,variable) %>% 
  summarize(mean = mean(mean), q5 = mean(q5), q95 = mean(q95)) %>% 
    ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_grid(subs~variable)+
  geom_pointrange()


params_icc %>% filter(grepl("ICC",variable) & residual_variance == T) %>% group_by(trials,variable) %>% 
  summarize(mean = mean(mean), q5 = mean(q5), q95 = mean(q95)) %>% 
    ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_grid(~variable)+
  geom_pointrange()



params_icc %>% filter(grepl("resid",variable)) %>% mutate(subs = as.factor(subs)) %>% 
  ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_wrap(~variable, nrow = 2, scales = "free")+
  geom_pointrange()

params_icc %>% filter(grepl("within",variable)) %>% mutate(subs = as.factor(subs)) %>% 
  ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_wrap(~variable, nrow = 2, scales = "free")+
  geom_pointrange()

params_icc %>% filter(grepl("ICC",variable), residual_variance == F) %>% mutate(subs = as.factor(subs)) %>% 
  ggplot(aes(y = mean, ymin = q5, ymax = q95, x = trials))+facet_wrap(~variable, nrow = 2, scales = "free")+
  geom_pointrange()

```


```{r, fig.height=6, fig.width=8}
trial <- map_dfr(results2, 2)
  
para = "betarts"
  
q = trial %>% filter(parameter == para) %>% dplyr::select(-c("resp","prob","X","trials.1","rts")) %>% distinct() %>% 
  pivot_longer(cols = c("real_beta_uncon","real_lambda_uncon","real_alpha_uncon","real_betarts_uncon","real_shift_uncon","real_sigma_uncon","real_inter_uncon")) %>% 
  filter(paste0("real_",para, "_uncon") == name) %>% filter(trials %in% c(40,120,200))

qq = q %>% ggplot(aes(x = value, y = mean, ymin = q5, ymax = q95, group = sim_id))+
  geom_pointrange()+facet_grid(n_subs~trials)+geom_abline(slope = 1, intercept = 0)

qq


para = "beta"
  
q = trial %>% filter(parameter == para) %>% dplyr::select(-c("resp","prob","X","trials.1","rts")) %>% distinct() %>% 
  pivot_longer(cols = c("real_beta_uncon","real_lambda_uncon","real_alpha_uncon","real_betarts_uncon","real_shift_uncon","real_sigma_uncon","real_inter_uncon")) %>% 
  filter(paste0("real_",para, "_uncon") == name) %>% filter(trials %in% c(40,120,200))

qq = q %>% ggplot(aes(x = value, y = mean, ymin = q5, ymax = q95, group = sim_id))+
  geom_pointrange()+facet_grid(n_subs~trials)+geom_abline(slope = 1, intercept = 0)

qq





qa = trial %>% filter(n_subs == 30, trials == 200)

trial%>% filter(n_subs == 30, trials == 200) %>% ggplot(aes(x = X, y = rts, col = as.factor(sessions))) +geom_point()+facet_wrap(~subs)+scale_x_continuous(limits = c(-40,40))


```




# scripts






fit_pathfinder_static_rts = function(parameters){
  
  N = parameters$trials
  
  
  r = array(NA,N)
  p = array(NA,N)
  x = array(NA,N)
  rts = array(NA,N)
  
  parm_ev = data.frame()
  
  x[1] = 0
  p[1] = parameters$real_lambda_con + (1 - 2 * parameters$real_lambda_con) * (0.5+0.5*erf((x[1]-parameters$real_alpha_con)/(parameters$real_beta_con*sqrt(2))))
  r[1] = rbinom(1,1,p[1])
  
  
  rts[1] = rlnorm(1,parameters$real_inter_con+parameters$real_betarts_con * (p[1]*(1-p[1])), parameters$real_sigma_con)+parameters$real_shift_con
  
  
  mod = cmdstanr::cmdstan_model(here::here("Stanmodels","pathfinder.stan"))
  
  for(i in 1:N){
    
    
    data_stan = list(trials = nrow((data.frame(r) %>% drop_na())),
                     x = data.frame(x) %>% drop_na() %>% .$x,
                     r = data.frame(r) %>% drop_na() %>% .$r)
    
    fit_psy <- mod$pathfinder(data = data_stan,refresh=0)
    
    if(i < 5){
      x[i] = data.frame(fit_psy$summary("psy_alpha")) %>% .$mean
    }else{
      sign = sample(c(-1,1),1)
      x[i] = fit_psy$draws("psy_alpha")[sample(1:1000,1)]+sign*fit_psy$draws("psy_beta")[sample(1:1000,1)]
    }
    
    if(abs(x[i]) > 50){
      print("to high")
      x[i] = data.frame(fit_psy$summary("psy_alpha")) %>% .$mean
    }
    
    
    c_trial = data.frame(fit_psy$summary(c("psy_alpha","psy_beta","psy_lambda"))) %>% mutate(trials = i)
    
    parm_ev = rbind(c_trial,parm_ev)
    p[i] = parameters$real_lambda_con + (1 - 2 * parameters$real_lambda_con) * (0.5+0.5*erf((x[i]-parameters$real_alpha_con)/(parameters$real_beta_con*sqrt(2))))
    r[i] = rbinom(1,1,p[i])
    
    rts[i] = rlnorm(1,parameters$real_inter_con+parameters$real_betarts_con * (p[i]*(1-p[i])), parameters$real_sigma_con)+parameters$real_shift_con
    
    if(rts[i] > 8){
      rts[i] = 8
    }
  }
  
  df_trial = data.frame(X = x,
                        prob = p,
                        resp = r,
                        rts = rts,
                        trials = parameters$trials,
                        real_lambda_con = parameters$real_lambda_con,
                        real_alpha_con = parameters$real_alpha_con,
                        real_beta_con = parameters$real_beta_con,
                        
                        real_inter_con = parameters$real_inter_con,
                        real_betarts_con = parameters$real_betarts_con,
                        real_sigma_con = parameters$real_sigma_con,
                        real_shift_con = parameters$real_shift_con,
                        
                        
                        real_lambda_uncon = parameters$real_lambda_uncon,
                        real_alpha_uncon = parameters$real_alpha_uncon,
                        real_beta_uncon = parameters$real_beta_uncon,
                        
                        real_inter_uncon = parameters$real_inter_uncon,
                        real_betarts_uncon = parameters$real_betarts_uncon,
                        real_sigma_uncon = parameters$real_sigma_uncon,
                        real_shift_uncon = parameters$real_shift_uncon,
                        
                        
                        id =  rnorm(1,0,1000),
                        subs = parameters$subs,
                        sessions = parameters$sessions,
                        model = "normal",
                        trials = 1:N)

  return(list(df_trial))
}




get_params = function(subs, trials){
  
  alpha_uncon = rnorm(subs,0,10)
  
  beta_uncon = rnorm(subs,3,0.6)
  
  lambda_uncon = rnorm(subs,-4,2)
  
  inter_uncon = rnorm(subs,-2,0.5)
  
  betarts_uncon = rnorm(subs,1,0.3)
  
  sigma_uncon = rnorm(subs,-1,0.3)
  
  shift_uncon = rnorm(subs,-1,0.5)
  
  replicate = 1:2
  
  parameters1 = NULL
  
  
  
  parameters = data.frame(real_alpha_uncon = alpha_uncon,
                          real_lambda_uncon = lambda_uncon,
                          real_beta_uncon = beta_uncon,
                          real_inter_uncon = inter_uncon,
                          real_betarts_uncon = betarts_uncon,
                          real_sigma_uncon = sigma_uncon,
                          real_shift_uncon = shift_uncon,
                          trials = trials)
  
  
  for(i in 1:(length(replicate))){
    parameters1 = rbind(parameters1,parameters %>% mutate(sessions = i))
  }
  
  parameters = parameters1 %>% 
    mutate(id = 1:nrow(.)) %>% 
    mutate(subs = rep(1:subs,2))
  
  parameters$real_beta_con = exp(parameters$real_beta_uncon)
  parameters$real_lambda_con = brms::inv_logit_scaled(parameters$real_lambda_uncon) / 2
  parameters$real_alpha_con = parameters$real_alpha_uncon
  
  parameters$real_inter_con = parameters$real_inter_uncon
  parameters$real_betarts_con = exp(parameters$real_betarts_uncon)
  parameters$real_sigma_con = exp(parameters$real_sigma_uncon)
  parameters$real_shift_con = exp(parameters$real_shift_uncon)
  
  
  
  data_list2 <- split(parameters, parameters$id)
  
  
  plan(multisession, workers = 4)
  
  #adding safety for if something goes wrong then it just outputs "Error" instead of crashing
  
  possfit_model = possibly(.f = fit_pathfinder_static_rts, otherwise = "Error")
  
  #test that it works
  #test  = possfit_model(data_list2[[1]])
  
  #run simulations!
  results <- future_map(data_list2, ~possfit_model(.x), .progress = TRUE, .options = furrr_options(seed = TRUE))
  
  
  trialwise_data = map_dfr(results, bind_rows)
  
  trialwise_data %>% ggplot(aes(x = X, y = rts))+geom_point()+facet_wrap(~subs)
  
  # write.csv(trialwise_data, here::here("datasets",paste0("subs = ", max(trialwise_data$subs),
  #                                                      " trials = ", trialwise_data$trials[1],
  #                                                      " id = ", rnorm(1,0,10),".csv")))
  # 
  # 
  return(trialwise_data)
}




get_ICC_psychometric = function(trialwise_data){
  
  sim_n_id = rnorm(1,0,10)
  
  #mod = cmdstanr::cmdstan_model(here::here("report","DDM","Stan Models","Psychometric_2_ICC.stan"))
  mod = cmdstanr::cmdstan_model(here::here("Stanmodels","Psycho_RTS_ICC_MSE.stan"))
  
  
  
  write.csv(trialwise_data, here::here("data","beta=3_rt_mean=1_sd=0.3",paste0("subs = ", max(trialwise_data$subs),
                                                                " trials = ", trialwise_data$trials[1],
                                                                " id = ", rnorm(1,0,10),".csv")))
  
  
  
  params = trialwise_data %>% arrange(sessions, subs)
  
  sesTsubs = length(unique(params$subs))*length(unique(params$sessions))
  
  data_stan = list(T = nrow(params %>% filter(subs == 1 & sessions == 1)),
                   P = length(unique(params$subs)),
                   S = length(unique(params$sessions)),
                   resp = matrix(params %>% arrange(sessions, subs) %>% .$resp, nrow = sesTsubs, byrow = T),
                   rts = matrix(params %>% arrange(sessions, subs) %>% .$rts, nrow = sesTsubs, byrow = T),
                   
                   X = matrix(params %>% arrange(sessions, subs) %>% .$X, nrow = sesTsubs, byrow = T),
                   
                   min_RT = params %>% group_by(sessions, subs) %>% summarize(minRT = min(rts)) %>% .$minRT,
                   
                   real_alpha = params %>% group_by(sessions, subs) %>% summarize(alpha = mean(real_alpha_uncon)) %>% .$alpha,
                   real_lambda = params %>% group_by(sessions, subs) %>% summarize(lambda = mean(real_lambda_uncon)) %>% .$lambda,
                   real_beta = params %>% group_by(sessions, subs) %>% summarize(beta = mean(real_beta_uncon)) %>% .$beta,
                   
                   real_inter = params %>% group_by(sessions, subs) %>% summarize(inter = mean(real_inter_uncon)) %>% .$inter,
                   real_betarts = params %>% group_by(sessions, subs) %>% summarize(beta_rt = mean(real_betarts_uncon)) %>% .$beta_rt,
                   real_sigma = params %>% group_by(sessions, subs) %>% summarize(sigma = mean(real_sigma_uncon)) %>% .$sigma,
                   real_shift = params %>% group_by(sessions, subs) %>% summarize(shift = mean(real_shift_uncon)) %>% .$shift
                   
                   
                   
  )
  
  #trialwise_data <- read_csv("datasets/subs = 14 trials = 90 id = 1.78080543202527.csv")
  # trialwise_data %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta")) %>% group_by(name) %>% summarize(mean = mean(value), sd = sd(value))
  # 
  # vector = fit$summary("beta") %>% .$mean
  # pairs <- lapply(1:(length(vector)/2), function(i) vector[c(i, i+14)])
  # pairwise_sd <- function(pair) {
  #   sd(pair)
  # }
  # 
  # # Calculate standard deviations of pairs
  # pairwise_sds <- unlist(lapply(pairs, pairwise_sd))
  # mean(pairwise_sds)
  
  fit <- mod$sample(
    data = data_stan,
    chains = 4,
    refresh = 50,
    init = 0,
    parallel_chains = 4,
    adapt_delta = 0.9,
    max_treedepth = 12)
  
  # rbind(
  #   fit$summary("alpha") %>% mutate(model = "no_RT"),
  #   fit_rt$summary("alpha") %>% mutate(model = "RT")
  # ) %>% ggplot(aes(x = rep(data_stan$real_alpha,2), y = mean, ymin = q5, ymax = q95, col = model))+geom_pointrange(position=position_dodge(width=2))+geom_abline()
  # 
  variances = rbind(data.frame(fit$summary("tau_b")) %>% 
                      mutate(variable = c("between_sub_var_beta","between_sub_var_lambda","between_sub_var_alpha",
                                          "between_sub_var_inter","between_sub_var_beta_rt","between_sub_var_sigma","between_sub_var_shift")),
                    data.frame(fit$summary("tau_w")) %>% 
                      mutate(variable = c("within_sub_var_beta","within_sub_var_lambda","within_sub_var_alpha",
                                          "within_sub_var_inter","within_sub_var_beta_rt","within_sub_var_sigma","within_sub_var_shift"))
  )
  
  
  residual = data.frame(fit$summary(c("resid_alpha_var","resid_beta_var","resid_lambda_var","resid_inter_var",
                                      "resid_betarts_var","resid_sigma_var","resid_shift_var")))
  
  extras = rbind(residual,variances)
  
  
  df = rbind(rbind(data.frame(fit$summary("ICC")) %>% mutate(variable = c("ICC_beta","ICC_lambda","ICC_alpha",
                                                                          "ICC_inter","ICC_beta_rt","ICC_sigma_rt","ICC_shift")) %>% 
                     mutate(residual_variance = F),
                   rbind(fit$summary("ICC_alpha"),fit$summary("ICC_lambda"),fit$summary("ICC_beta"),
                         fit$summary("ICC_inter"),fit$summary("ICC_beta_rt"),fit$summary("ICC_sigma"),fit$summary("ICC_shift"))%>% 
                     mutate(variable = c("ICC_alpha","ICC_lambda","ICC_beta",
                                         "ICC_inter","ICC_beta_rt","ICC_sigma","ICC_shift"),residual_variance = T)
  ),extras %>% mutate(residual_variance = NA)) %>% 
    mutate(diag = data.frame(fit$diagnostic_summary()) %>% summarize(mean_div = mean(num_divergent)),
           tree_depth = data.frame(fit$diagnostic_summary()) %>% summarize(mean_treedepth = mean(num_max_treedepth)),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  indi_estimates = rbind(data.frame(fit$summary(c("alpha","beta","lambda",
                                                  "inter","betarts","sigma","shift")))) %>% 
    mutate(id_model = gsub(".*?\\[(\\d+)\\].*?", "\\1", variable),
           parameter = gsub("\\[.*", "", variable),
           subs = rep(rep(1:length(unique(params$subs)), 7),2),
           sessions = rep(rep(1:2, each = length(unique(params$subs))),7),
           n_subs = length(unique(params$subs)),
           sim_id = sim_n_id)
  
  trial_estimates = inner_join(indi_estimates, params) %>% distinct()
  
  #correlations:
  correlation_no_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon",
                          "real_inter_uncon","real_betarts_uncon","real_sigma_uncon","real_shift_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = cor.test(mean, real_values)$estimate[1],
              correlation_q2 = cor.test(mean, real_values)$conf.int[[1]],
              correlation_q97 = cor.test(mean, real_values)$conf.int[[2]]
    )%>% mutate(ucnertainty_added = F)
  
  samples = 1000
  
  correlation_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon",
                          "real_inter_uncon","real_betarts_uncon","real_sigma_uncon","real_shift_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    rowwise() %>% 
    mutate(new_estimates = list(rnorm(samples, mean,sd)),
           id = list(1:samples)) %>%  unnest() %>% group_by(id, parameter) %>%
    summarize(correlation = cor.test(mean, new_estimates)$estimate[1]) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = mean(correlation),
              correlation_q2 = HDInterval::hdi(correlation)[[1]],
              correlation_q97 = HDInterval::hdi(correlation)[[2]]
    ) %>% mutate(ucnertainty_added = T)
  
  
  
  correlations = rbind(correlation_no_unc,correlation_unc) %>% 
    mutate(variable = paste0("ICC_",parameter),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  #plot:
  # trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta"), names_to = "real_parameternames",values_to = "real_values") %>% filter(paste0("real_",parameter) == real_parameternames) %>% ggplot(aes(x = mean, y = real_values))+geom_point()+facet_wrap(~parameter, scales = "free")
  rts = list(df, trial_estimates,correlations)
  
  
  mod = cmdstanr::cmdstan_model(here::here("Stanmodels","Psychometric_2_ICC_MSE.stan"))
  
  
  params = trialwise_data %>% arrange(sessions, subs)
  
  sesTsubs = length(unique(params$subs))*length(unique(params$sessions))
  
  data_stan = list(T = nrow(params %>% filter(subs == 1 & sessions == 1)),
                   P = length(unique(params$subs)),
                   S = length(unique(params$sessions)),
                   resp = matrix(params %>% arrange(sessions, subs) %>% .$resp, nrow = sesTsubs, byrow = T),
                   
                   X = matrix(params %>% arrange(sessions, subs) %>% .$X, nrow = sesTsubs, byrow = T),
                   
                   real_alpha = params %>% group_by(sessions, subs) %>% summarize(alpha = mean(real_alpha_uncon)) %>% .$alpha,
                   real_lambda = params %>% group_by(sessions, subs) %>% summarize(lambda = mean(real_lambda_uncon)) %>% .$lambda,
                   real_beta = params %>% group_by(sessions, subs) %>% summarize(beta = mean(real_beta_uncon)) %>% .$beta
                   
                   
  )
  
  #trialwise_data <- read_csv("datasets/subs = 14 trials = 90 id = 1.78080543202527.csv")
  # trialwise_data %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta")) %>% group_by(name) %>% summarize(mean = mean(value), sd = sd(value))
  # 
  # vector = fit$summary("beta") %>% .$mean
  # pairs <- lapply(1:(length(vector)/2), function(i) vector[c(i, i+14)])
  # pairwise_sd <- function(pair) {
  #   sd(pair)
  # }
  # 
  # # Calculate standard deviations of pairs
  # pairwise_sds <- unlist(lapply(pairs, pairwise_sd))
  # mean(pairwise_sds)
  
  fit <- mod$sample(
    data = data_stan,
    chains = 4,
    refresh = 50,
    init = 0,
    parallel_chains = 4,
    adapt_delta = 0.9,
    max_treedepth = 12)
  
  
  variances = rbind(data.frame(fit$summary("tau_b")) %>% 
                      mutate(variable = c("between_sub_var_beta","between_sub_var_lambda","between_sub_var_alpha")),
                    data.frame(fit$summary("tau_w")) %>% 
                      mutate(variable = c("within_sub_var_beta","within_sub_var_lambda","within_sub_var_alpha"))
  )
  
  
  residual = data.frame(fit$summary(c("resid_alpha_var","resid_beta_var","resid_lambda_var")))
  
  extras = rbind(residual,variances)
  
  
  df = rbind(rbind(data.frame(fit$summary("ICC")) %>% mutate(variable = c("ICC_beta","ICC_lambda","ICC_alpha")) %>% 
                     mutate(residual_variance = F),
                   rbind(fit$summary("ICC_alpha"),fit$summary("ICC_lambda"),fit$summary("ICC_beta"))%>% 
                     mutate(variable = c("ICC_alpha","ICC_lambda","ICC_beta"),residual_variance = T)
  ),extras %>% mutate(residual_variance = NA)) %>% 
    mutate(diag = data.frame(fit$diagnostic_summary()) %>% summarize(mean_div = mean(num_divergent)),
           tree_depth = data.frame(fit$diagnostic_summary()) %>% summarize(mean_treedepth = mean(num_max_treedepth)),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  indi_estimates = rbind(data.frame(fit$summary(c("alpha","beta","lambda")))) %>% 
    mutate(id_model = gsub(".*?\\[(\\d+)\\].*?", "\\1", variable),
           parameter = gsub("\\[.*", "", variable),
           subs = rep(rep(1:length(unique(params$subs)), 3),2),
           sessions = rep(rep(1:2, each = length(unique(params$subs))),3),
           n_subs = length(unique(params$subs)),
           sim_id = sim_n_id)
  
  trial_estimates = inner_join(indi_estimates, params) %>% distinct()
  
  #correlations:
  correlation_no_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = cor.test(mean, real_values)$estimate[1],
              correlation_q2 = cor.test(mean, real_values)$conf.int[[1]],
              correlation_q97 = cor.test(mean, real_values)$conf.int[[2]]
    )%>% mutate(ucnertainty_added = F)
  
  samples = 2000
  
  correlation_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    rowwise() %>% 
    mutate(new_estimates = list(rnorm(samples, mean,sd)),
           id = list(1:samples)) %>%  unnest() %>% group_by(id, parameter) %>%
    summarize(correlation = cor.test(mean, new_estimates)$estimate[1]) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = mean(correlation),
              correlation_q2 = HDInterval::hdi(correlation)[[1]],
              correlation_q97 = HDInterval::hdi(correlation)[[2]]
    ) %>% mutate(ucnertainty_added = T)
  
  
  
  correlations = rbind(correlation_no_unc,correlation_unc) %>% 
    mutate(variable = paste0("ICC_",parameter),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  
  
  norts = list(df, trial_estimates,correlations)
  
  
  
  return(list(rts,norts))
  
}


together = function(parameters){
  iccs = get_ICC_psychometric(get_params(parameters$subs,parameters$trials))
  return(list(iccs))
  
}








fit_without_data_gen= function(dav){
  
  sim_n_id = rnorm(1,0,10)
  
  #mod = cmdstanr::cmdstan_model(here::here("report","DDM","Stan Models","Psychometric_2_ICC.stan"))
  mod = cmdstanr::cmdstan_model(here::here("Stanmodels","Psycho_RTS_ICC_MSE.stan"))
  
  
  # 
  # write.csv(trialwise_data, here::here("data","beta=3 rts datasets",paste0("subs = ", max(trialwise_data$subs),
  #                                                                          " trials = ", trialwise_data$trials[1],
  #                                                                          " id = ", rnorm(1,0,10),".csv")))
  # 
  
  file  = list.files(here::here("data","beta=3 rts datasets"))[grepl("subs = 30",list.files(here::here("data","beta=3 rts datasets")))][dav$id]
  
  trialwise_data = read.csv(here::here("data","beta=3 rts datasets",file))
          
  params = trialwise_data %>% arrange(sessions, subs)
  
  sesTsubs = length(unique(params$subs))*length(unique(params$sessions))
  
  data_stan = list(T = nrow(params %>% filter(subs == 1 & sessions == 1)),
                   P = length(unique(params$subs)),
                   S = length(unique(params$sessions)),
                   resp = matrix(params %>% arrange(sessions, subs) %>% .$resp, nrow = sesTsubs, byrow = T),
                   rts = matrix(params %>% arrange(sessions, subs) %>% .$rts, nrow = sesTsubs, byrow = T),
                   
                   X = matrix(params %>% arrange(sessions, subs) %>% .$X, nrow = sesTsubs, byrow = T),
                   
                   min_RT = params %>% group_by(sessions, subs) %>% summarize(minRT = min(rts)) %>% .$minRT,
                   
                   real_alpha = params %>% group_by(sessions, subs) %>% summarize(alpha = mean(real_alpha_uncon)) %>% .$alpha,
                   real_lambda = params %>% group_by(sessions, subs) %>% summarize(lambda = mean(real_lambda_uncon)) %>% .$lambda,
                   real_beta = params %>% group_by(sessions, subs) %>% summarize(beta = mean(real_beta_uncon)) %>% .$beta,
                   
                   real_inter = params %>% group_by(sessions, subs) %>% summarize(inter = mean(real_inter_uncon)) %>% .$inter,
                   real_betarts = params %>% group_by(sessions, subs) %>% summarize(beta_rt = mean(real_betarts_uncon)) %>% .$beta_rt,
                   real_sigma = params %>% group_by(sessions, subs) %>% summarize(sigma = mean(real_sigma_uncon)) %>% .$sigma,
                   real_shift = params %>% group_by(sessions, subs) %>% summarize(shift = mean(real_shift_uncon)) %>% .$shift
                   
                   
                   
  )
  
  #trialwise_data <- read_csv("datasets/subs = 14 trials = 90 id = 1.78080543202527.csv")
  # trialwise_data %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta")) %>% group_by(name) %>% summarize(mean = mean(value), sd = sd(value))
  # 
  # vector = fit$summary("beta") %>% .$mean
  # pairs <- lapply(1:(length(vector)/2), function(i) vector[c(i, i+14)])
  # pairwise_sd <- function(pair) {
  #   sd(pair)
  # }
  # 
  # # Calculate standard deviations of pairs
  # pairwise_sds <- unlist(lapply(pairs, pairwise_sd))
  # mean(pairwise_sds)
  
  fit <- mod$sample(
    data = data_stan,
    chains = 4,
    refresh = 500,
    init = 0,
    parallel_chains = 4,
    adapt_delta = 0.9,
    max_treedepth = 12)
  
  # rbind(
  #   fit$summary("alpha") %>% mutate(model = "no_RT"),
  #   fit_rt$summary("alpha") %>% mutate(model = "RT")
  # ) %>% ggplot(aes(x = rep(data_stan$real_alpha,2), y = mean, ymin = q5, ymax = q95, col = model))+geom_pointrange(position=position_dodge(width=2))+geom_abline()
  # 
  variances = rbind(data.frame(fit$summary("tau_b")) %>% 
                      mutate(variable = c("between_sub_var_beta","between_sub_var_lambda","between_sub_var_alpha",
                                          "between_sub_var_inter","between_sub_var_beta_rt","between_sub_var_sigma","between_sub_var_shift")),
                    data.frame(fit$summary("tau_w")) %>% 
                      mutate(variable = c("within_sub_var_beta","within_sub_var_lambda","within_sub_var_alpha",
                                          "within_sub_var_inter","within_sub_var_beta_rt","within_sub_var_sigma","within_sub_var_shift"))
  )
  
  
  residual = data.frame(fit$summary(c("resid_alpha_var","resid_beta_var","resid_lambda_var","resid_inter_var",
                                      "resid_betarts_var","resid_sigma_var","resid_shift_var")))
  
  extras = rbind(residual,variances)
  
  
  df = rbind(rbind(data.frame(fit$summary("ICC")) %>% mutate(variable = c("ICC_beta","ICC_lambda","ICC_alpha",
                                                                          "ICC_inter","ICC_beta_rt","ICC_sigma_rt","ICC_shift")) %>% 
                     mutate(residual_variance = F),
                   rbind(fit$summary("ICC_alpha"),fit$summary("ICC_lambda"),fit$summary("ICC_beta"),
                         fit$summary("ICC_inter"),fit$summary("ICC_beta_rt"),fit$summary("ICC_sigma"),fit$summary("ICC_shift"))%>% 
                     mutate(variable = c("ICC_alpha","ICC_lambda","ICC_beta",
                                         "ICC_inter","ICC_beta_rt","ICC_sigma","ICC_shift"),residual_variance = T)
  ),extras %>% mutate(residual_variance = NA)) %>% 
    mutate(diag = data.frame(fit$diagnostic_summary()) %>% summarize(mean_div = mean(num_divergent)),
           tree_depth = data.frame(fit$diagnostic_summary()) %>% summarize(mean_treedepth = mean(num_max_treedepth)),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  indi_estimates = rbind(data.frame(fit$summary(c("alpha","beta","lambda",
                                                  "inter","betarts","sigma","shift")))) %>% 
    mutate(id_model = gsub(".*?\\[(\\d+)\\].*?", "\\1", variable),
           parameter = gsub("\\[.*", "", variable),
           subs = rep(rep(1:length(unique(params$subs)), 7),2),
           sessions = rep(rep(1:2, each = length(unique(params$subs))),7),
           n_subs = length(unique(params$subs)),
           sim_id = sim_n_id)
  
  trial_estimates = inner_join(indi_estimates, params) %>% distinct()
  
  #correlations:
  correlation_no_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon",
                          "real_inter_uncon","real_betarts_uncon","real_sigma_uncon","real_shift_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = cor.test(mean, real_values)$estimate[1],
              correlation_q2 = cor.test(mean, real_values)$conf.int[[1]],
              correlation_q97 = cor.test(mean, real_values)$conf.int[[2]]
    )%>% mutate(ucnertainty_added = F)
  
  samples = 1000
  
  correlation_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon",
                          "real_inter_uncon","real_betarts_uncon","real_sigma_uncon","real_shift_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    rowwise() %>% 
    mutate(new_estimates = list(rnorm(samples, mean,sd)),
           id = list(1:samples)) %>%  unnest() %>% group_by(id, parameter) %>%
    summarize(correlation = cor.test(mean, new_estimates)$estimate[1]) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = mean(correlation),
              correlation_q2 = HDInterval::hdi(correlation)[[1]],
              correlation_q97 = HDInterval::hdi(correlation)[[2]]
    ) %>% mutate(ucnertainty_added = T)
  
  
  
  correlations = rbind(correlation_no_unc,correlation_unc) %>% 
    mutate(variable = paste0("ICC_",parameter),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  #plot:
  # trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta"), names_to = "real_parameternames",values_to = "real_values") %>% filter(paste0("real_",parameter) == real_parameternames) %>% ggplot(aes(x = mean, y = real_values))+geom_point()+facet_wrap(~parameter, scales = "free")
  rts = list(df, trial_estimates,correlations)
  
  
  mod = cmdstanr::cmdstan_model(here::here("Stanmodels","Psychometric_2_ICC_MSE.stan"))
  
  
  params = trialwise_data %>% arrange(sessions, subs)
  
  sesTsubs = length(unique(params$subs))*length(unique(params$sessions))
  
  data_stan = list(T = nrow(params %>% filter(subs == 1 & sessions == 1)),
                   P = length(unique(params$subs)),
                   S = length(unique(params$sessions)),
                   resp = matrix(params %>% arrange(sessions, subs) %>% .$resp, nrow = sesTsubs, byrow = T),
                   
                   X = matrix(params %>% arrange(sessions, subs) %>% .$X, nrow = sesTsubs, byrow = T),
                   
                   real_alpha = params %>% group_by(sessions, subs) %>% summarize(alpha = mean(real_alpha_uncon)) %>% .$alpha,
                   real_lambda = params %>% group_by(sessions, subs) %>% summarize(lambda = mean(real_lambda_uncon)) %>% .$lambda,
                   real_beta = params %>% group_by(sessions, subs) %>% summarize(beta = mean(real_beta_uncon)) %>% .$beta
                   
                   
  )
  
  #trialwise_data <- read_csv("datasets/subs = 14 trials = 90 id = 1.78080543202527.csv")
  # trialwise_data %>% pivot_longer(cols = c("real_lambda","real_alpha","real_beta")) %>% group_by(name) %>% summarize(mean = mean(value), sd = sd(value))
  # 
  # vector = fit$summary("beta") %>% .$mean
  # pairs <- lapply(1:(length(vector)/2), function(i) vector[c(i, i+14)])
  # pairwise_sd <- function(pair) {
  #   sd(pair)
  # }
  # 
  # # Calculate standard deviations of pairs
  # pairwise_sds <- unlist(lapply(pairs, pairwise_sd))
  # mean(pairwise_sds)
  
  fit <- mod$sample(
    data = data_stan,
    chains = 4,
    refresh = 500,
    init = 0,
    parallel_chains = 4,
    adapt_delta = 0.9,
    max_treedepth = 12)
  
  
  variances = rbind(data.frame(fit$summary("tau_b")) %>% 
                      mutate(variable = c("between_sub_var_beta","between_sub_var_lambda","between_sub_var_alpha")),
                    data.frame(fit$summary("tau_w")) %>% 
                      mutate(variable = c("within_sub_var_beta","within_sub_var_lambda","within_sub_var_alpha"))
  )
  
  
  residual = data.frame(fit$summary(c("resid_alpha_var","resid_beta_var","resid_lambda_var")))
  
  extras = rbind(residual,variances)
  
  
  df = rbind(rbind(data.frame(fit$summary("ICC")) %>% mutate(variable = c("ICC_beta","ICC_lambda","ICC_alpha")) %>% 
                     mutate(residual_variance = F),
                   rbind(fit$summary("ICC_alpha"),fit$summary("ICC_lambda"),fit$summary("ICC_beta"))%>% 
                     mutate(variable = c("ICC_alpha","ICC_lambda","ICC_beta"),residual_variance = T)
  ),extras %>% mutate(residual_variance = NA)) %>% 
    mutate(diag = data.frame(fit$diagnostic_summary()) %>% summarize(mean_div = mean(num_divergent)),
           tree_depth = data.frame(fit$diagnostic_summary()) %>% summarize(mean_treedepth = mean(num_max_treedepth)),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  indi_estimates = rbind(data.frame(fit$summary(c("alpha","beta","lambda")))) %>% 
    mutate(id_model = gsub(".*?\\[(\\d+)\\].*?", "\\1", variable),
           parameter = gsub("\\[.*", "", variable),
           subs = rep(rep(1:length(unique(params$subs)), 3),2),
           sessions = rep(rep(1:2, each = length(unique(params$subs))),3),
           n_subs = length(unique(params$subs)),
           sim_id = sim_n_id)
  
  trial_estimates = inner_join(indi_estimates, params) %>% distinct()
  
  #correlations:
  correlation_no_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = cor.test(mean, real_values)$estimate[1],
              correlation_q2 = cor.test(mean, real_values)$conf.int[[1]],
              correlation_q97 = cor.test(mean, real_values)$conf.int[[2]]
    )%>% mutate(ucnertainty_added = F)
  
  samples = 2000
  
  correlation_unc = trial_estimates %>% dplyr::select(-c("X","resp","prob","trials.1")) %>% distinct() %>% 
    pivot_longer(cols = c("real_lambda_uncon","real_alpha_uncon","real_beta_uncon"), names_to = "real_parameternames",values_to = "real_values") %>% 
    filter(paste0("real_",parameter,"_uncon") == real_parameternames) %>% 
    rowwise() %>% 
    mutate(new_estimates = list(rnorm(samples, mean,sd)),
           id = list(1:samples)) %>%  unnest() %>% group_by(id, parameter) %>%
    summarize(correlation = cor.test(mean, new_estimates)$estimate[1]) %>% 
    group_by(parameter) %>% 
    summarize(correlation_mean = mean(correlation),
              correlation_q2 = HDInterval::hdi(correlation)[[1]],
              correlation_q97 = HDInterval::hdi(correlation)[[2]]
    ) %>% mutate(ucnertainty_added = T)
  
  
  
  correlations = rbind(correlation_no_unc,correlation_unc) %>% 
    mutate(variable = paste0("ICC_",parameter),
           trials = unique(params$trials),
           subs = length(unique(params$subs)),
           sessions = length(unique(params$sessions)),
           sim_id = sim_n_id)
  
  
  
  
  norts = list(df, trial_estimates,correlations)
  
  
  
  return(list(rts,norts))
  
}


